\bf On Devon Allen's Disqualification at the 2022 World Track and Field Championships

### 2024-11-12 Manuscript/manuscript.tex
started revisions

\begin{abstract}
Devon Allen was disqualified from the men's 110-meter hurdle final at
the 2022 World Track and Field Championships after recording a
reaction time of 0.099 seconds—just 0.001 seconds faster than the
allowable threshold of 0.1 seconds. This paper explores whether the
reaction times of athletes at the 2022 Championships were unusually
fast compared to other competitions and evaluates the appropriateness
of the 0.1-second disqualification barrier. To investigate these
issues, we employ a signed-rank test for clustered data to compare
athletes' reaction times across multiple competitions, along with a
generalized linear mixed model (GLMM) that accounts for both venue and
heat effects. The results indicate that athletes at the 2022
Championships had significantly faster reaction times on average. This
analysis raises important questions about the validity of the current
reaction time threshold and suggests potential modifications to the
rules that could prevent future disqualifications under similar
circumstances.

@@

The rest of this paper is organized as follows. Section~\ref{sec:Data} 
describes how the data were collected. The signed-rank test, detailed 
in Section~\ref{sec:rank}, uses slightly different data from the 
generalized linear mixed-effects model (GLMM) developed in 
Section~\ref{sec:glmm}. Section~\ref{sec:Results} presents the results 
of both analyses and discusses the implications of the reaction time 
barrier. Finally, Section~\ref{sec:concludingremarks} outlines the 
paper’s impact and limitations. All data and code for our analysis 
are provided in the supplementary materials.

@@

There are two types of data that are explored throughout this paper. For the
signed-rank tests described in Section~\ref{sec:rank}, the data comes from 
athletes who competed in the 2022 World Championships and another competition 
(2019 World Championships, 2023 World Championships, or 2022 national 
championships). For the GLMM described in Section~\ref{sec:glmm}, data comes
from every world championships from 1999 to 2023. This investigation started 
shortly after the 2022 World Championships, but as the 2023 data has now become 
available, we are pleased to include it in both of our analyses, and it did not
significantly change our results \citep{WAData}.

@@

\subsubsection{2022 National Competitions}\label{sec:datanational}
From June 23-26, 2022, the United States held its Track and Field Championships 
at Hayward Field in Eugene, Oregon to decide who to send to the World 
Championships being held at the same venue in August. Thus, we can establish a 
baseline for the four United States athletes who competed in at least one round 
of both events: Trey Cunningham, Daniel Roberts, Grant Holloway, and Devon Allen.  
Every athlete reacted faster in all the World Track and Field Championships 
races compared to the United States Track and Field Championships races. The 
dataset of the four United States athletes is too small to perform any 
meaningful statistical analysis, but the data was expanded to include reaction 
times for 11 more athletes who competed at other national competitions (many of which were 
held from May-July of 2022). This data was difficult to find, as the data is not
centrally located but instead found on different national websites (often 
recorded in the national language without translation).

Each athlete was designated as a ``cluster'' and the stage of competition (national
or international) was specified for each competition within the cluster (athlete).
Thus, this is subunit level grouping where we assign the treatment, national or
international, within each cluster. Cluster sizes range from three to six, with
one cluster of size three, six clusters of size four, two clusters of size five,
and two clusters of size six. We chose to include preliminary rounds in this
section as well as in Section~\ref{sec:data2019} in order to increase the size 
of our data. Without data from the preliminary rounds, the number of observations
dropped from 65 to 26 and the number of athletes included in the analysis
decreased from 15 to 9.

@@

While it is theoretically possible that athletes got faster 
between 2019 and 2022 or 2022 and 2023, it is important to 
remember that these are some of the fastest sprinters on the planet, and thus
improving on elite times is not easy. Once again we repeat a similar procedure
as in Section~\ref{sec:datanational}; we assign each athlete to be a cluster and
within each cluster label times to be from either the 2019, 2022, or 2023 World
Championships. When comparing 2019 and 2022 we look at data from 15
athletes, with two clusters of size two, three clusters of size three, five 
clusters of size four, two clusters of size five, and three clusters of size six.
When comparing 2022 and 2023 we look at data from 22 athletes, with seven 
clusters of size two, three clusters of size three, six clusters of size four,
and six clusters of size five.

\begin{figure}[tbp]
  \centering
  \includegraphics{RankScatterPlots}
  \caption{The top graph shows the reaction times of athletes who competed at 
	least once each at the 2019 and 2022 World Championships. The middle graph 
	shows the reaction times of athletes who competed at least once each at the 
	2022 and 2023 World Championships. The bottom graph shows the reaction times 
	of athletes who competed at least once at their country's
        2022 national championships and the 2022 World Championships.}
  \label{fig:RankScatterplots}
\end{figure}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics{Boxplot}
  \caption{The reaction times from 1999 to 2023 for the men's 110 meter hurdle.}
  \label{fig:Boxplot}
\end{figure}

>>>>

\begin{abstract}
Devon Allen was disqualified from the men's 110-meter hurdle final at
the 2022 World Track and Field Championships after recording a
reaction time of 0.099 seconds—just 0.001 seconds faster than the
allowable threshold of 0.1 seconds. This paper explores whether the
reaction times of athletes at the 2022 Championships were unusually
fast compared to other competitions and evaluates the appropriateness
of the 0.1-second disqualification barrier. To investigate these
issues, we employ a rank-sum test for clustered data to compare
athletes' reaction times across multiple competitions, along with a
generalized linear mixed model (GLMM) that accounts for both venue and
heat effects. The results indicate that athletes at the 2022
Championships had significantly faster reaction times on average. This
analysis raises important questions about the validity of the current
reaction time threshold and suggests potential modifications to the
rules that could prevent future disqualifications under similar
circumstances.

@@

The rest of this paper is organized as follows. Section~\ref{sec:Data} 
describes how the data were collected. The rank-sum test, detailed 
in Section~\ref{sec:rank}, uses slightly different data from the 
generalized linear mixed-effects model (GLMM) developed in 
Section~\ref{sec:glmm}. Section~\ref{sec:Results} presents the results 
of both analyses and discusses the implications of the reaction time 
barrier. Finally, Section~\ref{sec:concludingremarks} outlines the 
paper’s impact and limitations. All data and code for our analysis 
are provided in the supplementary materials.

@@

There are two types of data that are explored throughout this paper. For the
rank-sum tests described in Section~\ref{sec:rank}, the data comes from male 
athletes who competed in the 110 meter hurdles and 100 meter dash, and female
athletes who competed 100 meter hurdles and 100 meter dash.
Athletes must have competed in the 2022 World Championships and another competition 
(2019 World Championships, 2023 World Championships, or 2022 national 
championships). For the GLMM described in Section~\ref{sec:glmm}, data comes
from every world championships from 1999 to 2023 and includes reaction times from
the 110 meter hurdles and 100 meter dash. This investigation started 
shortly after the 2022 World Championships, but as the 2023 data has now become 
available, we are pleased to include it in both of our analyses, and it did not
significantly change our results \citep{WAData}.

@@

\subsubsection{2022 National Competitions}\label{sec:datanational}
From June 23-26, 2022, the United States held its Track and Field Championships 
at Hayward Field in Eugene, Oregon to decide who to send to the World 
Championships being held at the same venue in August. Thus, we can establish a 
baseline for the four United States athletes who competed in at least one round 
of both events: Trey Cunningham, Daniel Roberts, Grant Holloway, and Devon Allen.  
Every athlete reacted faster in all the World Track and Field Championships 
races compared to the United States Track and Field Championships races. The 
dataset of the four United States athletes is too small to perform any 
meaningful statistical analysis, but the data was expanded to include reaction 
times from more athletes who competed at other national competitions (many of which were 
held from May-July of 2022). This data was difficult to find, as the data is not
centrally located but instead found on different national websites (often 
recorded in the national language without translation).

Each athlete was designated as a ``cluster'' and the stage of competition (national
or international) was specified for each competition within the cluster (athlete).
Thus, this is subunit level grouping where we assign the treatment, national or
international, within each cluster. Cluster sizes range from three to six, with
one cluster of size three, six clusters of size four, two clusters of size five,
and two clusters of size six.
\of{This explanation is going to need to be re-worked now that we have more data.
Should we make a table of some sort? It's going to be very wordy without one}

We chose to include preliminary rounds in this section as well as in
Section~\ref{sec:data2019}, as removing them would have resulted in a much
narrower analysis with fewer athletes and clusters.

@@

While it is theoretically possible that athletes got faster 
between 2019 and 2022 or 2022 and 2023, it is important to 
remember that these are some of the fastest sprinters on the planet, and thus
improving on elite times is not easy. Once again we repeat a similar procedure
as in Section~\ref{sec:datanational}; we assign each athlete to be a cluster and
within each cluster label times to be from either the 2019, 2022, or 2023 World
Championships. When comparing 2019 and 2022 we look at data from 15
athletes, with two clusters of size two, three clusters of size three, five 
clusters of size four, two clusters of size five, and three clusters of size six.
When comparing 2022 and 2023 we look at data from 22 athletes, with seven 
clusters of size two, three clusters of size three, six clusters of size four,
and six clusters of size five.
\of{This explanation might need to change}

\begin{figure}[tbp]
  \centering
  \includegraphics{RankScatterPlots}
  \caption{The top graph shows the reaction times of athletes who competed at 
	least once each at the 2019 and 2022 World Championships. The middle graph 
	shows the reaction times of athletes who competed at least once each at the 
	2022 and 2023 World Championships. The bottom graph shows the reaction times 
	of athletes who competed at least once at their country's
  2022 national championships and the 2022 World Championships.}
  \label{fig:RankScatterplots}
\end{figure}
\of{Need to go through and edit this}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics{Boxplot}
  \caption{The reaction times from 1999 to 2023 for the men's 110 meter hurdle.}
  \label{fig:Boxplot}
\end{figure}
\of{Do We want 2 boxplot figures: 1 hurdles, 1 dash or combine them}
